Standardization of ceramic shape: A case studay from the Iron Age pottery from northeastern Taiwan

### 2020-08-13 analysis/paper/paper.Rmd
update chronology section to include modeling result

The archaeological record of Kiwulan's upper component shows traces of foreign contact, including Europeans in the 17th century, and waves of Chinese immigrants in the 19th century. Imported ceramics from mainland China, stoneware, and ornaments such as beads have been recovered in the upper component, indicating frequent long-distance trade activities with Europeans and Chinese merchants. Archaeological features such as burials, middens, and post-holes with _in-situ_ posts are widespread across the 1-2 m thick deposit of the upper component, and demonstrate that Kiwulan was a continuously occupied large settlement site [@Chen2007]. To compare different foreign influences, we classified the upper component into three chronological phases: pre-European, European, and Chinese. These phases were identified according to chronologically diagnostic artifacts based on their fined-grained time scale. Despite the presence of 17 radiocarbon ages [@Chen2007], they tend to have long time ranges that can only serve as a cross-validation if applicable. The diagnostic artifacts include blue and white porcelains, light grey glazed jars, and large dark brown glazed stoneware jars commonly used in the 17th century, and bricks and tiles employed by the Chinese in the 19th century [@Chen2007; @Hsieh2009; @Wang2011]. We also examined excavation depth measurement and stratigraphic details reported by the excavators (color, texture, disturbance, etc.) to reliably separate three phases. The deposit exhibited signs of continuous human occupation in each of the three phases with no apparent breaks. More details for the assignment of different phases are in the Online Supplementary Materials.

>>>>

The archaeological record of Kiwulan's upper component shows traces of foreign contact, including Europeans in the 17th century, and waves of Chinese immigrants in the 19th century. Imported ceramics from mainland China, stoneware, and ornaments such as beads have been recovered in the upper component, indicating frequent long-distance trade activities with Europeans and Chinese merchants. Archaeological features such as burials, middens, and post-holes with _in-situ_ posts are widespread across the 1-2 m thick deposit of the upper component, and demonstrate that Kiwulan was a continuously occupied large settlement site [@Chen2007]. To compare different foreign influences, we classified the upper component into three chronological phases: pre-European, European, and Chinese. These phases were identified according to chronologically diagnostic artifacts based on their fined-grained time scale. Despite the availability of 17 radiocarbon ages [@Chen2007], the chronological model demonstrates that time range are too coarse to provide sufficient information for the separation of three phases [@WangandMarwick2020]. Thus, radiocarbon ages only serve as a cross-validation for the diagnostic artifacts if applicable. The diagnostic artifacts include blue and white porcelains, light grey glazed jars, and large dark brown glazed stoneware jars commonly used in the 17th century, and bricks and tiles employed by the Chinese in the 19th century [@Chen2007; @Hsieh2009; @Wang2011]. We also examined excavation depth measurement and stratigraphic details reported by the excavators (color, texture, disturbance, etc.) to reliably separate three phases. The deposit exhibited signs of continuous human occupation in each of the three phases with no apparent breaks. More details for the assignment of different phases are in the Online Supplementary Materials.
